Difference-in-Differences with Spatial Spillovers

### 2021-04-07 latex/slides/Spillover_slides.tex
Update slides for Oral Comps

    \vspace{5mm}
    \textbf{Spillover effects} are when effect of treatment extend over the treatment boundaries such as states, counties, etc.

    \begin{itemize}
        \item e.g. large employer opening/closing in a \textbf{treated} county have positive employment effects on \textit{nearby counties}
    \end{itemize}
\end{frame}

@@

@@

\begin{frame}{Parallel Trends}
    I assume a modified version of the parallel counterfactual trends assumption: 
    \begin{align*}
        &\mathbb{E}\big[ \underbrace{Y_{i,1}(0, \textcolor{blue}{\vec{0}}) - Y_{i,0}(0, \textcolor{blue}{\vec{0}})}_{\text{Counterfactual Trend}} \mid D_i = 1 \big] \\
        = \ &\mathbb{E}\big[ \underbrace{Y_{i,1}(0, \textcolor{blue}{\vec{0}}) - Y_{i,0}(0, \textcolor{blue}{\vec{0}})}_{\text{Counterfactual Trend}} \mid D_i = 0 \big],
    \end{align*}

@@

\begin{frame}{What does Difference-in-Differences identify?}
    With the parallel trends assumption and random assignment of $D_i$, I decompose the diff-in-diff estimate as follows:

@@

\begin{frame}{Example of Bias: Large Factory Opening}\label{slide:example_bias}
    \textbf{Control Spillover:}

    $Y_{i1}(0, h_i(\vec{D})) - Y_{i1}(0, 0) = 1$ if county $i$ is within 40 miles of a treated county and $0$ otherwise.

    \textbf{Treated Spillover:}

    $Y_{i1}(1, h_i(\vec{D})) - Y_{i1}(1, 0) = 0.5$ if county $i$ is within 40 miles of a treated county and $0$ otherwise.

    \pause
    Then, the bias term is 
    \begin{align*}
        &0.5 * \text{\% of treated counties within 40 miles of another treated county} \\
        &-1 * \text{\% of treated counties within 40 miles of a treated county}
    \end{align*}

    \hyperlink{slide:partial_identification}{\beamerbutton{Partial Identification}}
\end{frame}

@@

\begin{frame}{Monte Carlo Simulations}
    \begin{itemize}
        \item Data Generating Process 
        \begin{align*}
            y_{it} &= \mu_t + \mu_i + \textcolor{green}{\tau_{\text{direct}}} D_{it} + \textcolor{purple}{\tau_{\text{spillover, control}}} (1-D_{it}) \text{Within 40mi.}_{it} \\ 
            &\quad + \textcolor{red}{\tau_{\text{spillover, treated}}} D_{it} \text{Within 40mi.}_{it} + \varepsilon_{it}
        \end{align*}

        \item Observation $i$ is a U.S. county, year $t \in \{2000, \dots, 2019\}$, treatment $D_{it}$ turns on in 2010 and is assigned randomly or through a spatial process.

        \item $\text{Within 40mi.}_{it}$ is an indicator for having your county center of population be within 40 miles of a treated unit.

        \item $\textcolor{green}{\tau_{\text{direct}}} = 2$, $\textcolor{purple}{\tau_{\text{spillover, control}}} = 1$ and $\textcolor{red}{\tau_{\text{spillover, treat}}} \in \{0, 0.5\}$.
    \end{itemize}
\end{frame}

\begin{frame}{Source of Bias 1: Control Units}
    Simulation 1:

    \begin{itemize}
        \item The first simulation changes the probability of treatment, $D_{it}$.

        \item Assume, for now, there is only spillover on the control, i.e. $\tau_{\text{spillover,treat}} = 0$.

        \item Estimate via TWFE: \begin{equation*}
            y_{it} = \mu_t + \mu_i + \tau D_{it} + \epsilon_{it}    
        \end{equation*}

        \item The bias of our estimate is \textit{Bias} $= \hat{\tau} - 2$
    \end{itemize}

\imageframe{../../figures/figure-bias_from_prob_treat_slides.png}

\begin{frame}{Solution: Removing ``contaminated'' controls}
    A common solution to the problem of spillover is to restimate on a subsample with neighboring control units removed.

    This simulation drops only units with $\text{Within 40mi.}_{it} = 1$. The exact exposure mapping is typically not observable for researchers, so this is best case scenario.
\end{frame}

\imageframe{../../figures/figure-bias_fix_slides.png}

\begin{frame}{Why Removing Controls Doesn't Work Well}
    \begin{itemize}
        \item[1.] Lose precision in treatment effect estimate by dropping control units
        \item[2.] If there are spillovers on treated units, removing controls won't remove both sources of bias. 
    \end{itemize}

    I repeat the same simulations but with $\textcolor{red}{\tau_{\text{spillover, treat}}} = 0.5$

\imageframe{../../figures/figure-bias_fix_treat_slides.png}

@@

    This is not observable to a researcher which raises the question:

@@

    Then, I estimate each data-generating process using (potentially) misspecified $\tilde{h}(\vec{D}, i)$ and report the average estimate bias.

@@

    \begin{itemize}
        \item \textit{Within $40/80$mi.}:
        \begin{itemize}
            \item Inidicator for nearest treated unit being within $40/80$ miles.
        \end{itemize}

@@

    \onslide<3->{
        \textbf{Inidicator that captures all affected unit removes all bias}
    }
\end{frame}

@@

    \begin{itemize}
        \item $y$ are outcomes for agricultural employment, manufacturing employment, and median family income.
        \item $\text{TVA}_c$ is the treatment variable
        \item $X_{c, 1940}$ allow for different long-term trends based on covariates in 1940. 
    \end{itemize}

@@

            \item Manufacturing jobs that would have been created in the control units in the absence of treatment might move to the Tennessee Valley (negative spillover on control units)
        \end{itemize}

        \item \textbf{Income:} 
        \begin{itemize}
            \item Since manufacutring jobs pay higher than agricultural jobs, the above two spillover effects can cause changes in median family income
        \end{itemize}
    \end{itemize}
\end{frame}

@@

\begin{frame}
\begin{table}[ht]
    \caption{Effects of Tennessee Valley Authority on Decadel Growth}
    \label{tab:tva}
    \renewcommand{\arraystretch}{1.2}

@@

                \onslide<2->{
                    \\
                    Manufacturing employment    & $0.0560^{***}$ &  $0.0461^{**}$ &    $-0.0104$   &    $-0.0128$   &  $-0.0248^{*}$ \\
                    &   $(0.0187)$   &   $(0.0210)$   &   $(0.0205)$   &   $(0.0257)$   &   $(0.0147)$
                }
                \onslide<3->{
                    \\
                    Median family income        & $0.0191^{***}$ &  $0.0196^{**}$ &    $0.0056$    &    $-0.0062$   &    $-0.0011$   \\
                    &   $(0.0067)$   &   $(0.0077)$   &   $(0.0079)$   &   $(0.0060)$   &   $(0.0032)$ 
                }

@@

% ------------------------------------------------------------------------------
\appendix\section{Appendix}
% ------------------------------------------------------------------------------

\begin{frame}{Partial Identification}\label{slide:partial_identification}
    Researchers can use the above method to create bounds for the biases resulting from spillovers. The choices should follow from the empirical context at hand.

    To formalize this, I follow the literature on partial identification for treatment effects
    \begin{citecolor}
        [\citet{Armstrong_Kolesár_2020}, \citet{Rambachan_Roth_2020}, and \citet{Manski_Pepper_2018}]
    \end{citecolor}
\end{frame}

\begin{frame}{Partially-identified Set}
    Following \citet{Rambachan_Roth_2020} and \citet{Manski_Pepper_2018}, researchers can provide bounds to the maximum spillovers.

    Let $\Delta$ denote the interval of potential biases: 
    \begin{align*}
        \Delta &\equiv [\underline{\Delta}, \overline{\Delta}] \\
        &= \left[ \min \tau_{\text{spill,treated}} - \tau_{\text{spill,control}}, \ \max \tau_{\text{spill,treated}} - \tau_{\text{spill,control}} \right] 
    \end{align*}

    The partially identified set is formed by \[ 
        \left[\hat{\tau} - \underline{\Delta} , \ \hat{\tau} + \overline{\Delta} \right]    
    \]

\begin{frame}{Bias-adjusted Inference}

    Inference can be done following \citet{Rambachan_Roth_2020} by creating a Fixed-Length Confidence Interval: 
    \[ 
        \left[
            \hat{\tau} - \underline{\Delta} - z_{1 - \alpha/2} \text{SE}_{\hat{\tau}}, \ \hat{\tau} + \overline{\Delta} + z_{1 - \alpha/2} \text{SE}_{\hat{\tau}}
        \right]
    \]

    These standard errors should account for the spatial structure of outcomes following \citet{Conley_1999}.

    \hyperlink{slide:example_bias}{\beamerbutton{Back to Example of Bias}}
\end{frame}

>>>>

    \vspace{5mm}
    \textbf{Spillover effects} are when effect of treatment extend over the treatment boundaries (states, counties, etc.). Example:

    \begin{itemize}
        \item A large employer opening/closing in a treated county have positive employment effects on \textbf{control counties}

        \item Having nearby counties with factories raises wages and reduces effect of \textbf{treated counties}
    \end{itemize}
\end{frame}

@@

\begin{frame}{Remove Bias}
    \only<1>{
        \[ y_{it} = \mu_t + \mu_i + \tau D_{it} + \varepsilon_{it} \]
    }
    \only<2>{
        \[ 
            y_{it} = \mu_t + \mu_i + \tau D_{it} + \text{Within}_{it} * D_{it} + \text{Within}_{it} * (1-D_{it}) + \varepsilon_{it}
        \]
    }

    \only<1>{ 
        \[ 
            \mathbb{E}\{\hat{\tau}\} = \tau + \text{\color{purple} Spillover on Control} + \text{\color{red}Spillover on Treated}
        \]
    }
    \only<2>{
        \[ 
            \mathbb{E}\{\hat{\tau}\} = \tau,
        \]
        so long as $\text{Within}_{it}$ contains all the units with spillovers.
    }
\end{frame}

@@

\begin{frame}{Parallel Trends}
    I assume a modified version of the parallel counterfactual trends assumption:

    \textbf{Assumption:} \textit{Parallel Counterfactual Trends}
    \begin{align*}
        &\mathbb{E}\big[ \underbrace{Y_{i,1}(0, \textcolor{blue}{\vec{0}}) - Y_{i,0}(0, \textcolor{blue}{\vec{0}})}_{\text{Counterfactual Trend}} \mid D_i = 1 \big] \\
        = \ &\mathbb{E}\big[ \underbrace{Y_{i,1}(0, \textcolor{blue}{\vec{0}}) - Y_{i,0}(0, \textcolor{blue}{\vec{0}})}_{\text{Counterfactual Trend}} \mid D_i = 0 \big],
    \end{align*}

@@

\begin{frame}{What does Difference-in-Differences identify?}
    With the parallel trends assumption and random assignment of $D_i$, I decompose the difference-in-differences estimate as follows:

@@

\begin{frame}{Removing Bias}
    \textbf{Assumption:} \textit{Spillovers are Local}

    Let $d(i,j)$ be the distance between units $i$ and $j$. There exists a distance $\bar{d}$ such that

        (i) For all units $i$,
        \[ 
            \min_{j: \ D_j = 1} d(i,j) > \bar{d} \implies h(\vec{D}, i) = \vec{0}. 
        \]

        (ii) There are treated and control units such that $\min_{j: \ D_j = 1} d(i,j) > \bar{d}$.
\end{frame}

\begin{frame}{Removing Bias}
    With assumption that spillovers are local, define $S_i$ to be an indicator that contains all units with $h(\vec{D}, i) \neq \vec{0}$ (and potentially some units with $= \vec{0}$).

    Then $\hat{\tau}$ will be a consistent estimate for $\tau$ in this specification: 
    \[ 
        y_{it} = \mu_t + \mu_i + \tau D_{it} + \text{Within}_{it} * D_{it} + \text{Within}_{it} * (1-D_{it}) + \varepsilon_{it}
    \]

@@

\begin{frame}{Common Solution: Removing ``contaminated'' controls}
    A common solution to the problem of spillover is to restimate on a subsample with neighboring control units removed.

    This is not recommended for two reasons:

    \begin{itemize}
        \item Removing control units from the sample decreases precision of the estimates

        \item Spillover effects on treated units will still remain in the estimand. 
    \end{itemize}
\end{frame}

@@

    The corret exposure mapping is not observable to a researcher which raises the question:

@@

    \begin{itemize}
        \item Observation $i$ is a U.S. county, year $t \in \{2000, \dots, 2019\}$, treatment $D_{it}$ turns on in 2010 and is assigned randomly.
    \end{itemize}

    Then, I estimate each data-generating process using (potentially) misspecified $\tilde{h}(\vec{D}, i)$ and report the average estimate bias.
\end{frame}

@@

    \begin{itemize}
        \item \textit{Within $40/80$mi.}:
        \begin{itemize}
            \item Indicator for nearest treated unit being within $40/80$ miles.
        \end{itemize}

@@

    \onslide<3->{
        \textbf{Indicator (or set of indicators) that captures all affected unit removes all bias}
    }
\end{frame}

@@

    \begin{itemize}
        \item $y$ are outcomes for agricultural employment and manufacturing employment.
        \item $\text{TVA}_c$ is the treatment variable
        \item $X_{c, 1940}$ allow for different long-term trends based on covariates in 1940. 
    \end{itemize}

@@

            \item Manufacturing jobs that would have been created in the control units in the absence of treatment might move to the Tennessee Valley (negative spillover on control units)
        \end{itemize}
    \end{itemize}
\end{frame}

@@

\begin{frame}
\begin{table}[ht]
    \caption{Effects of Tennessee Valley Authority on Decadel Growth, 1940-2000}
    \label{tab:tva}
    \renewcommand{\arraystretch}{1.2}

@@

                \onslide<2->{
                    \\
                    Manufacturing employment    & $0.0560^{***}$ &  $0.0461^{**}$ &    $-0.0104$   &    $-0.0128$   &  $-0.0248^{*}$ \\
                    &   $(0.0187)$   &   $(0.0210)$   &   $(0.0205)$   &   $(0.0257)$   &   $(0.0147)$
                }

@@

### 2021-04-07 latex/slides_short/Spillover_slides_short.tex
Update slides for Oral Comps

    \vspace{5mm}
    \textbf{Spillover effects} are when effect of treatment extend over the treatment boundaries (states, counties, etc.)

    \begin{itemize}
        \item E.g. a large employer opening/closing in a \textbf{treated} county have positive employment effects on \textit{nearby counties}

>>>>

    \vspace{5mm}
    \textbf{Spillover effects} are when effect of treatment extend over the treatment boundaries (states, counties, etc.). Example:

    \begin{itemize}
        \item A large employer opening/closing in a \textbf{treated} county have positive employment effects on \textit{nearby counties}
